Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-08-14 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to QPD chapter
14/08/25 changes, chp 5 done (ref needed)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation Details}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}). The beam 
is focused by a objective with numerical aperture of 1.2.
The polarisation is defined using the Jones vector which 
describes the magnitude of the x and y components of the 
electric field (so the vector $[1.0, 0.0i]$ describes an 
electric field that only oscillates in the x direction). 
Unless stated otherwise the Jones vector is kept at $[1.0, 
0.0i]$.

As a paradigmatic example, we study a simple spherical 
dimer composed of polystyrene, while there is the 
possibility of considering absorption we choose to keep 
the refractive index fixed at 1.59. The ratio of the two 
sphere radii is given by $a_{I}/a_{II}$ where $a_{I}$ is 
kept at $1\ \mu m$ unless specified otherwise. For 
simulations we assume that the dimer is suspended in water 
at room temperature ($T = 298K,\ \eta = 1.0x10^{-3}Pa\ s$).

In order to compute the magnitude of the Brownian forces we
need to know the dimer's diffusion tensor \textbf{D}. Unlike
for spheres where the translational and rotational diffusion 
coefficients are equal along all three Cartesian axis', a
dimer's translational and rotational diffusion is dependent
on whether we consider the its long or short axis. We use
the results from \cite{Nir1973} to compute the dimensionless 
diffusion tensor and then scale it according to the radius of
the largest sphere ($a_I$). The dimer's centre of diffusion
is located a distance of $a_I\zeta$ from the point of contact
between the two spheres, the coefficient $\zeta$ is based on
the size ratio of the two spheres. We use the centre of 
diffusion as the target origin that is provided to \textit{mstm}

@@

Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
to perform a systematic study of the dynamics demonstrated by 
asymmetric dimers in both plane and circularly polarised light. 
The simulations have two primary frames of reference that require 
explanation. Firstly, there is the laboratory frame which 
describes the coordinate system, we define the origin of our 
simulation as the focus of our trapping beam. While the choice in 
direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
specifically chosen so that it is aligned with the direction of 
beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation 
state of the particle, orientations are reported as quaternions which 
can be readily transformed into rotation matrices. The dimer's 
orientation is given by a rotation matrix in the form: 
\begin{align}
	\begin{pmatrix}
		U_{1,x} & U_{2,x} & U_{3,x} \\
		U_{1,y} & U_{2,y} & U_{3,y} \\
		U_{1,z} & U_{2,z} & U_{3,z}
	\end{pmatrix}
\end{align}
The matrix describes the direction of the particles 3 orientation 
axis $U_1$, $U_2$, and $U_3$. The latter is oriented so that it 
is collinear with the centres of the two spheres; the direction 
of vector $U_3$ is such that it goes from the centre of $a_{II}$ 
to the centre of $a_{I}$. The origin of the 3 axis' is set on the 
particle's centre of mass, this is also used to define it's 
displacement from the origin of the laboratory frame. The diffusion 
tensor of the dimer is more complex than a sphere's as we must 
consider the rotational effects experienced by the dimer. We use 
the analytical solution of Nir \& Acrivos \cite{Nir1973}, where 
they provided the solution coefficients depending on the dimer's 
size ratio ($a_I/a_{II}$). Rather than recalculate the coefficients 
for every possible configuration of dimer we use a spline fitting 
function to approximate the diffusion tensor.

We utilise a python library that utilises \textit{mstm} and
\textit{ott} to simulate the dynamical behaviour of spherical
aggregates. By specifying the optical properties of the target 
\textit{mstm} generates the targets T-matrix which can then be 
used by \textit{ott} to compute the optical forces generated by 
a focused beam. Using the Finite Differences approach we can solve
the Langevin Equation to take discrete time steps and update the 
dimer's position and orientation. A time step of $10 \mu s$ is 
used in line with previous works \cite{Vigilante2020}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation Details}
\subsection{Set up}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59$) located at the focus of a Gaussian 
beam. The dimer is composed of two spheres with radii 
$a_I$ and $a_{II}$, the size ratio between the two is 
given as $a_{II}/a_{I}$ - unless stated otherwise we 
kept $a_I$ fixed at $1\ \mu m$.

The beam is focused by a objective with numerical aperture 
of 1.2. The polarisation is defined using the Jones vector 
which describes the magnitude of the x and y components of 
the electric field (so the vector $[1.0, 0.0i]$ describes 
an electric field that only oscillates in the x direction). 
Unless stated otherwise the Jones vector is kept at $[1.0, 
0.0i]$. The conditions of the water are given as - $T = 298K,\ 
\eta = 1.0x10^{-3}Pa\ s,\ n_m = 1.33$.

\subsection{Coordinate system}
Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
to perform a systematic study of the dynamics demonstrated by 
asymmetric dimers in both plane and circularly polarised light. 
The simulations have two primary frames of reference that require 
explanation. Firstly, there is the laboratory frame which 
describes the coordinate system, we define the origin of our 
simulation as the focus of our trapping beam. While the choice in 
direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
specifically chosen so that it is aligned with the direction of 
beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation 
state of the particle, orientations are reported as quaternions which 
can be readily transformed into rotation matrices. The dimer's 
orientation is given by a rotation matrix in the form: 
\begin{align}
	\begin{pmatrix}
		U_{1,x} & U_{2,x} & U_{3,x} \\
		U_{1,y} & U_{2,y} & U_{3,y} \\
		U_{1,z} & U_{2,z} & U_{3,z}
	\end{pmatrix}
\end{align}
The matrix describes the direction of the particles 3 orientation 
axis $U_1$, $U_2$, and $U_3$. The latter is oriented so that it 
is collinear with the centres of the two spheres; the direction 
of vector $U_3$ is such that it goes from the centre of $a_{II}$ 
to the centre of $a_{I}$. The origin of the 3 axis' is set on the 
particle's centre of mass, this is also used to define it's 
displacement from the origin of the laboratory frame.

@@

\subsection{Diffusion Matrix}
In order to compute the magnitude of the Brownian forces we need to 
know the dimer's diffusion matrix \textbf{D}. Unlike for spheres 
where the translational and rotational diffusion coefficients are 
equal along all three Cartesian axis', a dimer's translational and 
rotational diffusion is dependent on whether we consider the its 
long or short axis. We use the results from \cite{Nir1973} to 
compute the dimensionless diffusion matrix and then scale it 
according to the radius of the largest sphere ($a_I$). The dimer's 
centre of diffusion is located a distance of $a_I\zeta$ from the 
point of contact between the two spheres, the coefficient $\zeta$ 
is based on the size ratio of the two spheres. We use the centre of 
diffusion as the target origin that is provided to \textit{mstm}

\subsection{Equations of Motion}
\label{sec:eq_of_motion}
The main equation of motion for an optically trapped particle is the 
Langevin equation. For a dimer the equation remains the same but is 
altered slightly to account for coupling between rotational and 
translational motion. 
\begin{align}
	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
	\sqrt{2\vec{D}_x}\eta(t)
\end{align}

Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
motion is generalised to a 3 dimensional case. Here we omit the inertial term 
as the restoration time of the optical trap is far greater than the characteristic time of the particle ($\kappa/\gamma \gg m/\gamma$). Now, the dimer is undergoing random rotational motion in addition to its Brownian 
translational motion the first term on the right hand side is no longer 
purely a function of the dimer's position but also of its orientation. 
The rotational form of the Langevin equation for a dipole within an external 
potential is given as \cite{Mazo2008}:
\begin{align}
	\frac{{d}\vec{u}(t)}{{dt}}
	=
	\frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
	+ \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}

Where $\vec{u}(t)$ is the unit vector aligned along the centres of 
the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the 
rotational friction coefficient which is given as $\gamma_R = 8\pi
\eta r^3$ for a sphere. The term $E(t)$ corresponds to the time 
dependent electric field vector, however we can assume that over a
time step as large as ours the time variation in the electric field 
is so insignificant that it can be treated as a constant. If we assume 
that over the course of a trajectory the dimer has a 'equilibrium' orientation in which the optical torque is minimised we can say that 
the first right hand term can be replaced with $\vec{\kappa}_u(\vec{r}
(t))$ to represent a harmonic potential due to the torque. $\lambda(t)$ 
is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian torque are normally distributed and are 
also uncorrelated so that:
\begin{align}
	\left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
\end{align}

For an asymmetric scatterer whose radius is comparable to that of the 
electric field's wavelength we now have a system of simultaneous equations:
\begin{align}
	\label{eq:full_langevin}
	\frac{{d}\vec{r}(t)}{{dt}}
	&=
	\frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2k_BT\vec{D}}\eta(t)
	\\
	\frac{{d}\vec{u}(t)}{{dt}}
	&=
	\frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
	+ \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}

Fortunately, we do not need to solve these directly as the latter two
random variables can be easily approximated if the thermal energy of
the system is known, and the rate of change can be assumed as linear if
we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma 
\ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
to compute the optical force and torque applied to the dimer.

\subsection{Computational methods}
Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
or \textit{ott} \cite{Nieminen2007} we can compute the beam 
shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
type. Using the results from \cite{Farsund1996} we can then 
compute both the optical force and torque using the beam 
coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
which are found by calculating the dimer's $T$-matrix via \textit{mstm} 
\cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
total force in along the z-direction and the total torque about 
the z-axis are provided:
\begin{equation}
	\label{eq:optical_force}
	\begin{split}
		\bold{F_z}
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
	\end{split}
\end{equation}
\begin{equation}
	\begin{split}
		\bold{T_z}
		&=
		-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
		\\ 
		&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
	\end{split}
\end{equation}

where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam 
coefficients of the incident and scattered fields respectively. 
We can get the $x$ and $y$ force and torque components in a 
similar form by applying a simple rotation transformation. With 
the optical forces and torques computed all that remains is to 
compute the Brownian forces and torques which are constrained 
by the relation.
\begin{align}
	\langle q_iq_j\rangle =2D_{ij}\Delta t
\end{align}

Vigilante et al compiled together a python package that combines 
both \textit{ott} and \textit{mstm} to simulate the behaviour of 
spherical aggregates within a predefined optical trap 
\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
modules that can be used to model the scattering from different 
shaped particles; however, for aggregates of spheres the resultant 
force is nonsensical (see Chapter 4 for further details). As such 
we rely on the \textit{mstm} for simulating dimers.

### 2025-08-14 Latex_files/Thesis/chapters/chapter 4/main.tex
14/08/25 changes, chp 5 done (ref needed)

\section{Interpretation of scattering data into orientation estimates}
We now discuss the methodology for computing the instantaneous 
orientation of an optically trapped particle based on scattering
detected by a finite grouping of optical fibres.

>>>>

\section{Interpretation of scattering data into orientation estimates}
\label{sec:ref_orientations_explained}
We now discuss the methodology for computing the instantaneous 
orientation of an optically trapped particle based on scattering
detected by a finite grouping of optical fibres.

### 2025-08-14 Latex_files/Thesis/chapters/chapter 5/main.tex
14/08/25 changes, chp 5 done (ref needed)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterisation of asymmetric dimer dynamics via PSD analysis}
While it is possible to extract all of the relevant dynamics from a simulation, from an experimental standpoint characterising those 
dynamics is dependent on what experimental techniques are used. 
Translational motion can be characterised via a position detection 
system but angular motion is far more difficult to detect, let along characterise.

The simulative QPD is composed of 4 photodiodes that measure the 
intensity of light incident on their surfaces. Given that the 
condensing lens will have a maximum acceptance angle given by 
its numerical aperture ($\theta_{max} = \sin^{-1}\left(NA_c/n_
{med}\right)$) we define the electric field incident on the QPD 
surface as the elements of the scattered and incident field that 
propagate within that angular range of $\pm \theta_{max}$. Using 
\textit{ott} we can evaluate the total electric field to an 
arbitrary spacial resolution across the surface of the QPD. The 
reported signal is found by taking the difference between pairs 
of quadrant signals (see \ref{sec:simulated_QPD} for further 
detail). This does not provide a one-to-one result however as 
hardware errors (i.e. internal resistance, external light sources, 
and vibrations) would distort the signal. It can be used however to 
show how if the reported power spectra is an accurate representation 
of the true particle dynamics or if the additional degrees of freedom 
have an adverse effect on the scattering. See chapter 2 for a breakdown
of power spectral analysis.

@@

\begin{center}
	\captionof{table}{QPD fitting for single sphere}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
			fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline

@@

This coupling of the translational and orientational motion 
has been highlighted previously \cite{Vigilante2020}, but 
there effects have not been demonstrated in the context of
an experimental situation. We can see from the difference in
trajectory and QPD fittings that while our description of the
trap shape is very similar (both methods give similar ellipticity
values) but the magnitude of the trap strength are significantly
different. This is partially explained by the fact that the
QPD is not actually measuring the position of the particle
but instead the intensity distribution of the total field 
incident on the QPD surface. In which case if the dimer is 
rotated slightly the QPD signal will change even if the dimers
centre of diffusion remains stationary.

\subsection{QPD for angular displacement detection}
The next logical step is to consider whether or not a QPD can at 
all be utilised for detection of rotational motion. This has been
shown for nano-particles that exhibit periodic rotational motion
using by considering the difference in diagonal quadrants 
\cite{Yifat2021}. The only difference between these results is 
the fact that rotation occurs perpendicular to the particle's long 
axis, whereas dimers rotate about their long axis. The benefit of 
detecting rotational motion is that one can begin to build a better understanding about how angular momentum is transferred to a dimer 
which could be extended to more complex particles. This could give 
some understanding to the results of Chapter 4.

@@

However if we compare this to fig.~\ref{fig:rohrbach} we can see 
that the the signal change is dwarfed by the translational motion. 
So while their is a clear relationship between the two trying to
discern between translational and rotational contributions to the
QPD signal is not a simple task. This is significant because it 
means that with regards to the power spectrum fittings in table 
\ref{tab:dimer} there is no way of ensuring that the measured 
trap strengths are accurate. The variance $\sigma(f_c)$ is only 
the variance based upon our fitting parameters and not on the 
actual dynamics of the dimer. In the case where the particle's
shape is not known prior its not possible to determine whether 
any inaccuracies in the power spectra are due to the fit or due
to the particle's motion.

The dependence of $S_x$ and $S_y$ and the dimer's orientation
(defined by the spherical angles $\theta$, $\phi$) is therefore
not clear. To that end, we utilised machine vector regression,
which takes as an input the voltage from the 4 quadrants and 
tries to fit that to the dimer's trajectory. Fitting the QPD
signal to the dimer's position shows promising results, shown in
fig~\ref{fig:QPD_rotation} is a prediction of a symmetric dimer's 
position based on the QPD signal. With the actual displacement 
on the x-axis and the predicted result y-axis, and the dashed 
line represents the y=x line - indicating an ideal prediction.

Fig~\ref{fig:QPD_rotation} shows that its relatively trivial 
to predict a particle's position based off the QPD signal, even 
in axial direction which has not been done previously. It should 
be noted that these displacements are relative to the trap focus 
and so accurate tracking of the beam movements needs to be taken. 
This result is partially due to the fact that displacement and 
signal units are linearly related by \eqref{eq:conversion_factor}. Interestingly when we run the same protocol for different sized 
dimer's there seems to be a cut off point when the fit begins to 
fail. Plotting the $R^2$ for each fit vs its size ratio shows 
that beyond a size ratio of 5 the machine learning begins to fall 
off significantly.

This is surprising because we would expect that as the second 
sphere shrinks the dynamics should more closely approximate that 
of a single sphere. This should be reflected in the total field
incident on the QPD and therefore the machine learning program 
should be able to easily detect the dimer's translational motion.
We can therefore only conclude that the dimer's rotational motion
is contributing to the scattering to the scattered field, even if
it is miniscule the contribution is enough to throw off a 
position detection system. Detection of rotational motion was 
attempted using machine learning, the model was trained on the 
same trajectory mentioned in \ref{sec:off-axis}, where the dimer 
is trapped in an off-axis orientation, from a top down perspective
it should be easy enough to note that the dimer is not vertically 
trapped even if the exact angle is difficult to determine.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterisation of asymmetric dimer dynamics via PSD analysis}
\label{sec:PSD_analysis}
While it is possible to extract all of the relevant dynamics from a simulation, from an experimental standpoint characterising those 
dynamics is dependent on what experimental techniques are used. 
Translational motion can be characterised via a position detection 
system but angular motion is far more difficult to detect, let along characterise.

The tweezer setup we are simulating incorporates a dual objective 
configuration, using one objective lens to focus the beam (NA=1.25)
where as the condenser lens acts as a means of truncating the electric
field ($NA_c=0.25$). The electric field falling onto the QPD surface is computed by considering the far field of the combined incident and 
scattered beams. We can ignore any contributions from the electric
field that falls outwith the maximum acceptance angle of our condenser
lens ($\theta_{max}\approx\pm10.8\deg$). This is not perfectly analogous
of a true QPD as external factors (i.e. internal resistance, external
light sources, and vibrations) would also contribute to the detected QPD 
signal. It does provide us a means of showing what elements of a particle's
trajectory are represented in the power spectrum and which elements are 
obscured.

@@

\begin{center}
	\captionof{table}{QPD fitting for single sphere}
	\label{tab:sphere}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
			fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline

@@

There is still a concern in the fact that both produce similar
power spectra, enough so that it would be difficult to determine
if a trapped particle was actually a dimer or a sphere. This is
because the rotational and translational motion, coupling of the 
has been highlighted previously \cite{Vigilante2020}, but there 
effects have not been demonstrated in the context of an experimental situation. We can see from the difference in trajectory and QPD 
fittings that while our description of the trap shape is very 
similar (both methods give similar ellipticity values) but the 
magnitude of the trap strength are significantly different. This 
is partially explained by the fact that the QPD is not actually 
measuring the position of the particle but instead the intensity 
distribution of the total field incident on the QPD surface. In 
which case if the dimer is rotated slightly the QPD signal will 
change even if the dimers centre of diffusion remains stationary. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{QPD for angular displacement detection}
\label{sec:angular_displacement}
The next logical step is to consider whether or not a QPD can at 
all be utilised for detection of rotational motion. This has been
shown for nano-particles that exhibit periodic rotational motion
using by considering the difference in diagonal quadrants 
\cite{Yifat2021}. The only difference between these results is 
the fact that rotation occurs perpendicular to the particle's long 
axis, whereas dimers rotate about their long axis. The benefit of 
detecting rotational motion is that one can begin to build a better understanding about how angular momentum is transferred to a dimer 
which could be extended to more complex particles. This could give 
some understanding to the results of Chapter 4.

@@

However if we compare this to fig.~\ref{fig:Rohrbach} we can see 
that the the signal change is dwarfed by the translational motion. 
So while their is a clear relationship between the two trying to
discern between translational and rotational contributions to the
QPD signal is not a simple task. This is significant because it 
means that with regards to the power spectrum fittings in Table 
\ref{tab:dimer} there is no means of discerning between that of 
a trapped sphere and that of a trapped dimer. Furthermore the 
dynamics reported by the power spectrum fitting do not accurately 
reflect the actual dynamics of the dimer, instead only providing 
a orientation-averaged trap strength. But unless we have some means
of at least estimating the degree of rotational freedom, calculating
the actual trapping strength is impossible.

The dependence of $S_x$ and $S_y$ and the dimer's orientation
(defined by the spherical angles $\theta$, $\phi$) is therefore
not clear. To that end, we utilised machine vector regression,
which takes as an input the voltage from the 4 quadrants and 
tries to fit that to the dimer's trajectory. We utilised the 
scikit-learn module available in Python to transform the 4D vector
of the QPD signal to a 2D vector specifying the expected displacement
of the target particle. Unlike in Chapter 4 where we utilised 
machine learning to get a probabilistic estimation of the 
particle's orientation we utilise vector regression as we 
want to replicate QPD precision as closely as possible.

In order to fit the QPD signal to dimer displacements we need
to train our model. For each size ratio considered a 3 second
trajectory was ran (using the same conditions in 
Sec.~\ref{sec:PSD_analysis}) and a normal distribution of the 
dimer's position was taken and used to construct a training 
data set of QPD signals. After the training results reached an
error rate of less than 1\% the same trajectory was used to
construct a test set that would allow our model to fit the QPD 
signal to the positional data from the trajectory. Fitting the 
QPD signal to the dimer's position showed promising results, 
shown in fig~\ref{fig:QPD_rotation} is a prediction of a 
symmetric dimer's position based on the QPD signal. With the 
actual displacement on the x-axis and the predicted result y-axis, 
and the dashed line represents the y=x line - indicating an 
ideal prediction. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{predicted_displacements.png}
	\caption{Scikit-Learn's best predictions of the displacement 
		of a symmetric dimer placed in an optical trap, predictions 
		are based on the QPD signal collected at the 4 quadrants. 
		(Left) predictions of the $X$ \& $Y$ displacements. 
		(Right) predictions of the $Z$ displacement, predictions 
		have been scaled so they represent the displacement from 
		the beam focus. The dashed black line represents when the 
		predicted displacement = the true displacement.}
	\label{fig:predicted_displacements}
\end{figure}

Fig~\ref{fig:predicted_displacements} shows that its relatively 
trivial to predict a particle's position based off the QPD signal
in the $X$-$Y$ plane but harder along the axial direction. If we 
consider Fig~\ref{fig:Rohrbach} we note that the change in signal
strength is only truly linear close to the beam focus, as shown in
Chapter 3, dimers do not trap close to the beam focus but further 
away. This would suggest that the model cannot accurately predict 
the displacement of a dimer in the axial direction as there is a 
non-linear relationship between the change in signal and the change 
in axial position. It should be noted that even in the worst case 
the difference between the predicted and true displacements is 
around $~20 nm$.

These displacements are relative to the trap focus and 
so accurate tracking of the beam movements needs to be taken if 
you want to consider the displacement relative to some larger object. 
These results are likely due to the fact that displacement and 
signal units are linearly related by \eqref{eq:conversion_factor}. Interestingly when we run the same protocol for different sized 
dimer's there seems to be a cut off point when the fit begins to 
fail. We can compute the coefficient of determination ($R^2$) for 
our model's prediction, the $R^2$ for each fit vs its size ratio 
shows that beyond a size ratio of 5 the machine learning begins 
to fall off significantly.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{score_results.png}
	\caption{The coefficient of determination for different sized 
	dimers using the scikit-learn vector regression model. Each line
	plot shows the $R^2$ value for either the $X$, $Y$, or $Z$ 
	displacements. When the size ratio is greater than 5 the $R^2$ 
	value falls off dramatically, from around 0.95 down to as low as 
	0.01.}
\end{figure}

This is surprising because we would expect that as the second 
sphere shrinks the dynamics should more closely approximate that 
of a single sphere. This should be reflected in the total field
incident on the QPD and therefore the machine learning program 
should be able to easily detect the dimer's translational motion.
We can therefore only conclude that the contribution that the 
second sphere makes to the scattered beam is not proportional to
its size. What is interesting is that the axial prediction does 
not experience as a large a drop in accuracy, suggesting that the
total field incident on the QPD is not changing drastically, but
rather that the second sphere is altering the distribution such
that the $X-Y$ displacements are being incorrectly predicted. 
This could simply be a limitation of the \textit{mstm} software being 
unable to accurately represent the T-matrix of the second sphere, 
however the documentation does not specify any limits on the relative 
size differences so it is unclear if this is a known issue. The 
problem is that it is difficult to verify the error between the 
\textit{mstm} predicted scattering and the true scattering without 
replicating the exact system experimentally. For now we can say that predictions for size ratios greater than 5 should be for the most 
part disregarded.

Detection of rotational motion was attempted using the same method, 
the model was trained on the same trajectory mentioned in 
\ref{sec:off-axis}, where the dimer is trapped in an off-axis 
orientation. We chose this trajectory instead of a strictly vertical
orientation in order to confirm that the model can distinguish between 
the polar angle and the azimuthal angle. The results of which are 
are shown in Fig.\ref{fig:predicted_rotations}
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{angular_predictions.png}
	\caption{Scikit-learn's prediction of the sphericical angles of
	a dimer trapped in an off-axis configuration (see Sec\ref{sec:off-axis}).
	(Left) predictions of the azimuthal angle, (Right) predictions 
	of the Polar angle. Dashed black lines show the ideal prerformance 
	where the predicted angle = true angle.}
	\label{fig:predicted_rotations}
\end{figure}

From here we can see that the vector regression model performs rather
poorly in predicting the rotations made by the dimer, even in a situation
where visibly the dimer is an off-axis orientation. While the model can 
roughly predict the median polar angle (~$102\deg$) it has no means of
predicting the rotations made by the dimer. This result is reveals two important factors: Firstly, the change in the QPD signal for rotational
motion is a full order of magnitude smaller than the change in signal 
for translational motion meaning the model has no means of distinguishing
between large and small rotations. Secondly, is the fact that unlike 
translational motion, rotational motion is dependent on the prior 
orientations in the trajectory. This was already discussed in Chapter 4
where we improved our machine learning model by accounting for the 
the priors $p(\hat{\bf n}_\alpha)$, where we weighted our estimation 
by the distance from our prior estimate. Vector regression is powerful
for predicting uncorrelated events but has no means for accounting for
cases where one event is dependent on the next. At the same time, 
utilising machine learning in the case of a QPD would not be able to 
provide a high enough orientational resolution without first sampling 
from a trajectory and creating a unique set of reference orientations
(see Sec.~ \ref{sec:ref_orientations_explained} for definitions).

@@

\section{Conclusion}
We have developed a program that allows one to compute the QPD 
response signal produced by an arbitrary shaped particle. By 
utilising \textit{mstm} we can construct any combination of 
spherical particles and measure the total field incident on the
focal plane of the condenser lens. This has immediate applications 
in replicating power spectral data from complex particle structures 
that experience periodic rotational motion. Furthermore, if all 
contributions to the QPD response signal are known it would allow 
researchers to compare the reported dynamics given by an experimental
QPD to that of our simulated response signal.

The QPD module shows promising results for dimers who have a size 
ratio less than 5. Utilising vector regression models allowed for 
accurate predictions of the displacements made by dimers while 
confined by an optical trap. The use of vector regression is necessary
to return a continuos range of positions instead of a probabilistic
approximation of the particle's position vector.

As shown in \ref{sec:angular_displacement} random rotational motion 
is difficult to distinguish from random translational motion. Even
when utilising a vector regression model the angular displacement
predicted does not match the trajectory data. This means that at least
for now utilising focal plane interferometry cannot provide an 
accurate estimation of angular displacements. While static light 
scattering methods - as seen in Chapter 4 - do present a possible 
alternative, the limited angular resolution due to a high mixing 
of signal and orientation spaces also make the method difficult to 
implement. Measurements of angular displacement are crucial for 
measuring torques exerted on trapped particles, current torque 
measurements assume that only the the polarisation contributes to 
the applied torque, measuring non-polarisation related torques still 
eludes us.

Overall, tracking the angular displacements of non-spherical 
objects for the purposes of torque measurements will require 
further study. Fully understanding the torques applied to an 
arbitrary shaped particle would pave the way for high-precision 
particle-manipulation techniques and a greater understanding 
of the interactions between focused beams and aggregate particles.
